\documentclass{article}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}

\input{defs-OCDL-LRU}
\begin{document}
Consider the problem:
\begin{equation}
\min_{\vx} \frac{\mu}{2} \|\mW^T\mQ\mW\vx - \mW^T\mQ\mW\vs\|_2^2 + \frac{\lambda}{2}\sum_{i} \|\mG_i\vx\|_2^2 
\end{equation}

where $\mQ$ is a nonlinear quantization operator such that $\vy = \mQ\vx$ implies $\vy[i] = \vq[i]*\operatorname{round}(\frac{\vx[i]}{\vq[i]})$.

I can rewrite this problem, splitting $\vx$ into two variables:
\begin{equation}
\begin{aligned}
\arg\min_{\vx,\vz} & \frac{\mu}{2} \|\vz - \vx\|_2^2 + \frac{\lambda}{2}\sum_{i} \|\mG_i\vx\|_2^2 \\
\text{subject to } & \mQ\mW\vz - \mQ\mW\vs = 0
\end{aligned} 
\end{equation}

It would certainly be possible to swap the constraint and the first objective term, but I would like to minimize the number of update equations for which the nonlinear operator $\mQ$ complicates.

The alternating directionn method of multipliers algorithm (ADMM) is designed for affine equality constrains, but $\mW^T\mQ\mW$ is close enough to linear that I expect it will succeed here despite the nonlinearity of quantization.

\section{$\vx$ Update}




\section{$\vz$ Update}
\begin{equation}
\vz^{(k + 1)} = \arg\min_{\vz} \frac{\mu}{2} \|\vz - \vx^{(k + 1)}\|_2^2 + \frac{\rho}{2}\|\mW^T\mQ\mW\vz + (1 - \alpha)\mW^T\mQ\mW\vz^{(k)} - (2 - \alpha)\mW^T\mQ\mW\vs + \frac{\vgamma^{(k)}}{\rho}\|_2^2
\end{equation}

I should note here that due to the quantization, this minimizer is not guarenteed (or even likely) to exist.  However, the infimum does exist and a solution that evaluates arbitrarily close to that infimum can be found.  

To solve this, I will start by ignoring the quantization and finding the assymptotic solution as $\vq$ approaches zero.
\begin{equation}
\vz^{(k + 1)} \approx \arg\min_{\vz} \frac{\mu}{2} \|\vz - \vx^{(k + 1)}\|_2^2 + \frac{\rho}{2}\|\mW^T\mW\vz + (1 - \alpha)\mW^T\mW\vz^{(k)} - (2 - \alpha)\mW^T\mW\vs + \frac{\vgamma^{(k)}}{\rho}\|_2^2
\end{equation}

\begin{equation}
(\mu\mId + \rho\mW^T\mW)\vz^{(k + 1)} \approx \mu\vx^{(k + 1)} + \rho\mW^T((2 - \alpha)\mW\vs - (1 - \alpha)\mW\vz^{(k)} - \frac{\vgamma}{\rho})
\end{equation}

\begin{equation}
\mu\mId + \rho\mW^T\mW = \mu(\mId - \mW^T\mW) + (\mu + \rho)\mW^T\mW
\end{equation}

\begin{equation}
(\mu\mId + \rho\mW^T\mW)^{-1} = \frac{1}{\mu}(\mId - \mW^T\mW) + \frac{1}{\mu + \rho}\mW^T\mW
\end{equation}

\begin{equation}
(\mu\mId + \rho\mW^T\mW)^{-1} = \frac{1}{\mu}\mId - \frac{\rho}{\mu^2 + \rho\mu}\mW^T\mW
\end{equation}

\begin{equation}
\vz^{(k + 1)} \approx \vx^{(k + 1)} - \frac{\rho}{\mu + \rho}\mW^T\mW\vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\mW^T((2 - \alpha)\mW\vs - (1 - \alpha)\mW\vz^{(k)} - \frac{\vgamma}{\rho})
\end{equation}

\begin{equation}
\vz^{(k + 1)} \approx \vx^{(k + 1)}  + \frac{\rho}{\mu + \rho}\mW^T((2 - \alpha)\mW\vs - \mW\vx^{(k + 1)} - (1 - \alpha)\mW\vz^{(k)} - \frac{\vgamma}{\rho})
\end{equation}

Now, I will reinsert $\mQ$ in the places it was removed. The result should still be fairly close to the actual solution.
\begin{equation}
\vz^{(k + 1)} \approx \vx^{(k + 1)}  + \frac{\rho}{\mu + \rho}\mW^T((2 - \alpha)\mQ\mW\vs - \mW\vx^{(k + 1)} - (1 - \alpha)\mQ\mW\vz^{(k)} - \frac{\vgamma}{\rho})
\end{equation}

In the interest of brevity, I will define:
\begin{equation}
\vr = (2 - \alpha)\mQ\mW\vs - \mW\vx^{(k + 1)} - (1 - \alpha)\mQ\mW\vz^{(k)} - \frac{\vgamma}{\rho}
\end{equation}

\begin{equation}
\vz_{\text{approx}}^{(k + 1)} = \vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\mW^T\vr
\end{equation}

Plugging the $\vz = \vz_{\text{approx}}^{(k + 1)} + \Delta\vz$ into the function to be minimized, I have the expression:
\begin{equation}
\frac{\mu}{2} \|\frac{\rho}{\mu + \rho}\mW^T\vr + \Delta\vz\|_2^2 + \frac{\rho}{2}\|\mQ\mW(\vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\mW^T\vr + \Delta\vz) - (2 - \alpha)\mQ\mW\vs + (1 - \alpha)\mQ\mW\vz^{(k)} + \frac{\vgamma^{(k)}}{\rho}\|_2^2
\end{equation}


\begin{equation}
\frac{\mu}{2} \|\frac{\rho}{\mu + \rho}\mW^T\vr + \Delta\vz\|_2^2 + \frac{\rho}{2}\|\mQ\mW(\vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\mW^T\vr + \Delta\vz) - \vr - \mW\vx^{(k + 1)}\|_2^2
\end{equation}

\begin{equation}
\mW\mW^T = \mId
\end{equation}

\begin{equation}
\frac{\mu}{2} \|\frac{\rho}{\mu + \rho}\mW^T\vr + \Delta\vz\|_2^2 + \frac{\rho}{2}\|\mQ(\mW\vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\vr + \mW\Delta\vz) - \vr - \mW\vx^{(k + 1)}\|_2^2
\end{equation}

Any component of $\Delta\vz$ that is orthogonal to $\mW$ will increase the first term of the objective expression and leave the second term unchanged. Therefore, the optimal choice of $\Delta\vz$ has no component orthogonal to $\mW$.

So, I can minimize this expression instead:

\begin{equation}
\frac{\mu}{2} \|\frac{\rho}{\mu + \rho}\vr + \mW\Delta\vz\|_2^2 + \frac{\rho}{2}\|\mQ(\mW\vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\vr + \mW\Delta\vz) - \vr - \mW\vx^{(k + 1)}\|_2^2
\end{equation}

Finally, again for convenience, define

\begin{equation}
\vy = \mW\vx^{(k + 1)} + \frac{\rho}{\mu + \rho}\vr
\end{equation}

So, I have
\begin{equation}
\frac{\mu}{2} \|\frac{\rho}{\mu + \rho}\vr + \mW\Delta\vz\|_2^2 + \frac{\rho}{2}\|-\frac{\mu}{\mu + \rho}\vr + \mQ(\vy + \mW\Delta\vz) - \vy\|_2^2
\end{equation}


It is convenient that when trying to minimize the expression, each element of $\mW\Delta\vz$ can be treated independently.

In comparison to $\mW\Delta\vz = 0$, it is possible to decrease the first objective term while not changing the second objective term by adjusting $\mW\Delta\vz$ in the direction $-\vr$.

\begin{equation}
\mW\Delta\vz[i] = -\operatorname{sign}(\vr[i])\operatorname{minimum}(|\frac{\rho}{\mu + \rho}\vr[i]|,|\vq[i]*\operatorname{round}(\frac{\vy[i]}{\vq[i]}) - 0.5^{-}\vq[i]\operatorname{sign}(\vr[i]) - \vy[i]|
\end{equation}
where $0.5^{-}$ is a number less than $0.5$, but arbitarily close to it.

In comparison to $\mW\Delta\vz = 0$, it may be possible to decrease the second term by using $\mW\Delta\vz$ to switch the rounding direction.
\begin{equation}
\mW\Delta\vz[i] = \vq[i]*\operatorname{round}(\frac{\vy[i]}{\vq[i]}) - 0.5^{+}\vq[i]\operatorname{sign}(\vy[i] - \vq[i]*\operatorname{round}(\frac{\vy[i]}{\vq[i]}) - \frac{\rho}{\mu + \rho}\vr[i]) - \vy[i]
\end{equation}

Finally, in comparison to $\mW\Delta\vz = 0$, it may be possible to decrease the first objective term while switching the rounding direction.
\begin{equation}
\mW\Delta\vz[i] = -\operatorname{sign}(\vr[i])\operatorname{minimum}(|\frac{\rho}{\mu + \rho}\vr[i]|,|\vq[i]*\operatorname{round}(\frac{\vy[i]}{\vq[i]}) - 1.5^{-}\vq[i]\operatorname{sign}(\vr[i]) - \vy[i]|
\end{equation}
By comparing these three possible minimizers\footnote{Again, I should clarify, not an actual minimizer, but something that evaluates arbitarily close to the infimum.}, I can identify the minimizer, and adjust the $\vz_{\text{approx}}^{(k + 1)}$ accordingly.
\begin{equation}
\vz^{(k + 1)} = \vz_{\text{approx}}^{(k + 1)} + \Delta\vz
\end{equation}

\section{$\vgamma$ update}

\begin{equation}
\frac{\vgamma^{(k + 1)}}{\rho} = \frac{\vgamma^{(k)}}{\rho} + \mQ\mW\vz^{(k + 1)} + (1 - \alpha)\mQ\mW\vz^{(k)} - (2 - \alpha)\mQ\mW\vs 
\end{equation}

\end{document}
